Fitting metabolic models to dissolved oxygen data: The Estuarine BAyesian Single-station Estimation method

### 2023-07-26 manu-draft.qmd
more revisions

$$
H\frac{dC}{dt} = aPAR - R - \left[bU_{10}^2\left(\frac{Sc}{600} \right)^{-0.5} \left(C_{Sat} - C \right )\right]
$$ {#eq-ebase}

where the terms on the right side of the equation are gross primary production (P or $aPAR$), ecosystem respiration (R), and gas exchange (D, remainder), respectively, modeled as a function of the change in dissolved oxygen ($dC$) per unit time ($dt$) of the input data. Production is modeled linearly as a function of surface irradiance ($PAR$, Watts m$^{-2}$) and $a$ (light efficiency, $\left(\text{mmol}/\text{m}^2/\text{d}\right)/\left(\text{W}/\text{m}^2\right)$). Respiration is modeled with the single parameter $R$ ($\text{mmol}/\text{m}^3/\text{d}$). Gas exchange is modeled using the formulation in @Wanninkhof14 as a wind-based parameterization for the gas transfer velocity [@Ho16] multiplied by the oxygen concentration gradient as the difference between the concentration measured in the surface water layer, $C$ (mmol m$^{-3}$), and the saturation (or equilibrium) concentration, $C_{Sat}$ (mmol m$^{-3}$), calculated from salinity (psu) and temperature ($^\circ$C) [@Millero81; @Garcia92]. The gas transfer velocity is modeled using wind speed squared at 10 meters above the water ($U_{10}^2$, m$^2$ s$^{-2}$), water column depth ($H$ in meters, as a fixed parameter or time series from a pressure sensor), and the Schmidt number ($Sc$) defined as the ratio of the kinematic viscosity of water to the molecular diffusivity of the gas for oxygen [from water temperature and salinity using the polynomial fit in @Wanninkhof14]. The $b$ parameter is estimated from the data and has units of $\left(\text{cm}/\text{hr}\right)/\left(\text{m}^2/\text{s}^2\right)$. As such, the three parameters estimated by EBASE are $a$, $R$, and $b$ using the required input data that includes time series of dissolved oxygen, water temperature, salinity, PAR, wind speed, and water column depth (`r run_reference(id = "tbl-ebaseiotab")`).

>>>>

$$
H\frac{dC}{dt} = aPAR - R - k_w\left(C_{Sat} - C \right )
$$ {#eq-ebase}

The parameter $k_w$ is the gas transfer velocity defined as:

$$
k_w = bU_{10}^2\left(\frac{Sc}{600} \right)^{-0.5}
$$ {#eq-kw}

For @eq-ebase, the terms on the right side of the equation are gross primary production (P, $aPAR$), ecosystem respiration (R), and gas exchange (D, remainder), respectively, processes that result in the change in dissolved oxygen ($dC$) per unit time ($dt$); $H$ is the water depth. Production is modeled linearly as a function of surface irradiance ($PAR$, $\text{W}/\text{m}^2$) and $a$ (light efficiency, $\left(\text{mmol}/\text{m}^2/\text{d}\right)/\left(\text{W}/\text{m}^2\right)$). Respiration is modeled with the single parameter $R$ ($\text{mmol}/\text{m}^2/\text{d}$). Gas exchange is modeled using the formulation in @Wanninkhof14 as a wind-based parameterization for the gas transfer velocity [@Ho16] multiplied by the oxygen concentration gradient as the difference between the concentration measured in the surface water layer, $C$ ($\text{mmol}/\text{m}^3$) , and the saturation (or equilibrium) concentration, $C_{Sat}$ ($\text{mmol}/\text{m}^3$) calculated from salinity (psu) and temperature ($^\circ$C) [@Millero81; @Garcia92]. The gas transfer velocity, $k_w$ in @eq-kw, is modeled using wind speed squared at 10 meters above the water ($U_{10}^2$, $\text{m}^2/\text{s}^2$) and the Schmidt number ($Sc$), which is the ratio of the kinematic viscosity of water to the molecular diffusivity of oxygen in water [$S_c$ is computed from water temperature and salinity using the polynomial fit in @Wanninkhof14]. The $b$ parameter is estimated from the data and has units of $\left(\text{cm}/\text{hr}\right)/\left(\text{m}^2/\text{s}^2\right)$. As such, the three parameters estimated by EBASE are $a$, $R$, and $b$ using the required input data that includes time series of dissolved oxygen, water temperature, salinity, PAR, wind speed, and water column depth (`r run_reference(id = "tbl-ebaseiotab")`).
